---
title: "Rule Learner - Classification"
output: 
  html_document: 
    highlight: pygments
    theme: journal
    toc: yes
    toc_float: yes
---


# Introduction:  

  * Rule learner algorithms generate if-else like rules to classify data based on feature values.  
  * Similar to *Decision Trees* this class of algorithms are also greedy - meaning they process the data first seen.  
  * The output is relatively easy to understand without statistical backgorund.  
  * *Decision Trees* divide and conquer while these seperate and conquer.  
  * Rule learners are computationally expensive compared to *Decision Trees*.  
  * Rule generators work best where features are factors.  
  * Similar to *Decision Trees* these also classify output based on feature values to increase homogeneity.  

# Application on 'Poison Mushroom' dataset.  
Below details copied from thw UCI website.  
Attribute Information: (classes: edible=e, poisonous=p)  
1.cap-shape:bell=b,conical=c,convex=x,flat=f,knobbed=k,sunken=s
2.cap-surface:fibrous=f,grooves=g,scaly=y,smooth=s
3.cap-color:brown=n,buff=b,cinnamon=c,gray=g,green=r,pink=p,purple=u,red=e,white=w,yellow=y
4.bruises?:bruises=t,no=f
5.odor:almond=a,anise=l,creosote=c,fishy=y,foul=f,musty=m,none=n,pungent=p,spicy=s
6.gill-attachment:attached=a,descending=d,free=f,notched=n
7.gill-spacing:close=c,crowded=w,distant=d
8.gill-size:broad=b,narrow=n
9.gill-color:black=k,brown=n,buff=b,chocolate=h,gray=g,green=r,orange=o,pink=p,purple=u,red=e,white=w,yellow=y  
10.stalk-shape:enlarging=e,tapering=t
11.stalk-root:bulbous=b,club=c,cup=u,equal=e,rhizomorphs=z,rooted=r,missing=?
12.stalk-surface-above-ring:fibrous=f,scaly=y,silky=k,smooth=s
13.stalk-surface-below-ring:fibrous=f,scaly=y,silky=k,smooth=s
14.stalk-color-above-ring:brown=n,buff=b,cinnamon=c,gray=g,orange=o,pink=p,red=e,white=w,yellow=y  
15.stalk-color-below-ring:brown=n,buff=b,cinnamon=c,gray=g,orange=o,pink=p,red=e,white=w,yellow=y  
16.veil-type:partial=p,universal=u
17.veil-color:brown=n,orange=o,white=w,yellow=y
18.ring-number:none=n,one=o,two=t
19.ring-type:cobwebby=c,evanescent=e,flaring=f,large=l,none=n,pendant=p,sheathing=s,zone=z  
20.spore-print-color:black=k,brown=n,buff=b,chocolate=h,green=r,orange=o,purple=u,white=w,yellow=y  
21.population:abundant=a,clustered=c,numerous=n, scattered=s,several=v,solitary=y  
22.habitat:grasses=g,leaves=l,meadows=m,paths=p,urban=u,waste=w,woods=d  

## Data Load:  

Read data and rename columns.

```{r RL_mus_load,  cache=T, warning=T}
df_raw <- read.csv('.//Data//agaricus-lepiota_data.csv',header = F, sep = ",")
colnames(df_raw) <- c('edi_psn',
'cap_shape',
'cap_surface',
'cap_color',
'bruises',
'odor',
'gill_attachment',
'gill_spacing',
'gill_size',
'gill_color',
'stalk_shape',
'stalk_root',
'stalk_surface_above_r',
'stalk_surface_below_r',
'stalk_color_above_rin',
'stalk_color_below_rin',
'veil_type',
'veil_color',
'ring_number',
'ring_type',
'spore_print_color',
'population',
'habitat'
)
dim(df_raw)
summary(df_raw)
df_raw <- subset(df_raw,select = c(colnames(df_raw)!='veil_type')) # this column has only 1 value so no infomation
dim(df_raw)
```
## Missing value  

Missing values are represented by '?'.  

```{r rl_mus_miss, cache=T, warning=T}
require(ggplot2)
require(gridExtra)
p1 <- ggplot(data=df_raw,aes(x=df_raw$stalk_root)) +
  geom_bar(stat='count',aes(fill=df_raw$edi_psn),position = 'dodge',width=0.6) +
  xlab('Stalk-root') +
  ylab('Mushroom-count') +
  ggtitle('Missing Values') +
  theme_grey()

p2 <- ggplot(data=df_raw,aes(x=df_raw$edi_psn)) +
  geom_bar(stat='count',aes(fill=df_raw$stalk_root),position = 'dodge',width=0.6) +
  xlab('Edible/Poison') +
  ylab('Mushroom-count') +
  ggtitle('Missing Values') +
  theme_grey()

grid.arrange(p1,p2,ncol=1)
```
  
- More than 25% of stalk-root are missing.  
- off these missing values 1/3 are edible and 2/3 are poisonous.  
- Other values of this variable have differing proportions edible/poisonous. 

Let's run the rule learner as is. If the rules include these '?' missing values let's think of imputing in correct way.  

## Rule Learner.  

```{r rl_mus_rulearn, cache=T,warning=T}
require(RWeka)
rl_mod <- JRip(edi_psn~.,data=df_raw) # RIPPER algorthm  implemented
summary(rl_mod)
rl_mod
```  

- These 9 rules correctly identifies all mushromms from the dataset.  
- They don't include 'Stalk-Root' variable which means the variable is not significant.
- No need to impute missing values.  

## Pair plot of variables included in 9 rules. 

All the rules have atmost 2 variables. So below set of pair graphs can be used as visual tool to cross verify the rules validity.  

```{r,fig.width=12,fig.height=20}
require(GGally)
require(ggplot2)
ggpairs(data=df_raw[,c('edi_psn',
'odor',
'gill_size',
'gill_color',
'spore_print_color',
'stalk_surface_below_r',
'stalk_surface_above_r',
'habitat',
'cap_color',
'stalk_color_above_rin')],
aes(colour=edi_psn),
columns=2:10,
title='Variable included in 9 rules')
```

## Rerun the JRip  

Some columns referred in the previous run. Cross check the output after running only these columns selected.

```{r rl_rerun, cache=T, warning=T}
df_spe <- df_raw[,c('edi_psn',
'odor',
'gill_size',
'gill_color',
'spore_print_color',
'stalk_surface_below_r',
'stalk_surface_above_r',
'habitat',
'cap_color',
'stalk_color_above_rin')]
rs_mod2 <- JRip(edi_psn~.,data=df_spe)
summary(rs_mod2)
rs_mod2
```

  
These 2 out puts exactly match and classify all mushrooms correctly.  

## Predict.  

Standard 90/10 train,  test split to verify the performance of JRip rule learner algorithm in predicting the correct category of mushrooms.  

```{r rl_pred, cache=T,warning=T}
set.seed(20181204)
trainIndex <- sample(seq(1:nrow(df_raw)),nrow(df_raw)*0.9)
df_train <- df_raw[trainIndex,]
df_test <- df_raw[-trainIndex,]
mod_p1 <- JRip(edi_psn~.,data=df_train)
prd1 <- predict(mod_p1,df_test, type='class')
table(df_test$edi_psn,prd1)
```

It correctly predicted all edible and poison mushrooms.!!!